Developing a preprint culture in biology

### 2012-09-28 ms.tex
review the text

% No abstract needed.

@@

Preprints gained popularity 20 years ago with the advent of arXiv, an
open preprint server widely used in physics and mathematics
\cite{gin11}. Preprints are also integral to the culture of other
scientific fields.  Paul Krugman noted that, in economics, the
\emph{traditional model of submit, get refereed, publish, and then
  people will read your work broke down a long time ago. In fact, it
  had more or less fallen apart by the early 80s} \cite{kru12}. In
addition to a section in arXiv, economists have also the RePEc
(Research Papers in Economics) initiative, which aims to create an
archive of working papers, manuscripts, and book chapters.  Despite
the success of this approach in other fields, most manuscripts in
biology are never submitted to any public archive.
% JJA: This is a tenuous statement: MANY if not most biomedical labs
% in the US are funded by NIH, which requires submission to Pubmed
% Central, which many would consider a preprint archive.  In other
% words, a biomedical editor might take issue with this statement,
% especially if he considers "biology" to be just his part of biology
% --- I know plenty of people who only think of "biology" as
% biomedical science
In this article, we will first highlight the advantages of open
preprint servers for both scientists and publishers. We will also
debunk a few misconceptions, discuss the policies of major publishers
in biology, and briefly review the most popular open preprint servers.

\section{The case for open preprints}

The first and most often discussed advantage of arXiv and open
preprints is speed (Figure~\ref{fig:map}). The time between submission
and the official publication of a manuscript can be measured in
months, sometime in years. For all this time, the research is known
only to a select few: colleagues, editors, reviewers.  Thus, the
science cannot be used, discussed, or reviewed by the wider scientific
community.  This stage of manuscript preparation is a problem for both
scientists and publishers.  Manuscripts that are unknown cannot be
used and thus take more time to be cited. It has been shown that
high-energy physics, with its high ArXiv submission rate, had the
highest immediacy among physics and mathematics \cite{pra05}.
% JJA: Can you say what "immediacy" means?

@@

Public preprint servers offer a much fairer way to establish
intellectual priority by making the work available when done. Some
manuscripts will spend much more time in the review process than
others. Surprisingly, there is a perception in biology that public
preprints make it easier to steal ideas, as if a scientific idea was
only complete once published in a peer-reviewed journal.
Mathematicians and physicists have embraced arXiv in part to establish
priority in a fair way \cite{cal12}.
% JJA: I snipped the part about getting "snooped" (do you mean
% "scooped"?) as it seems normative and might annoy some readers.  The
% "should" should be implicit.

@@

Submitting to a preprint server is becoming more common in biology,
even though it still involves a minority of papers. The quantitative
biology section in arXiv is experiencing faster growth in submissions
than any other field \cite{cal12}.  Examples include a recent series
of papers on the theory of natural selection that was posted to ArXiv
simultaneously with its publication in the \emph{Journal of
  Evolutionary Biology}
\cite{JEB:JEB2431,JEB:JEB2498,JEB:JEB2378,JEB:JEB2373}.  Other
submissions in this category include evolutionary and ecological
theory by authors trained in physics and computer science.  Since
authors in these fields regularly check ArXiv, submitting preprints
may be the most effective way biologists can help others avoid
``repeated work'' \cite{de2011contribution} and form a synthetic
community of evolutionary theorists from disparate backgrounds.

Most scientific journals are preprint-friendly: Nature, PLOS, BMC,
PNAS, Science (mostly) \ref{table:policies}, and all the journals from
Elsevier and Springer.  The Ecological Society of America and the
Genetics Society of America recently changed their policies to allow
public preprints.  Few scientific publications will not consider a
manuscript submitted to a public preprint server.  Still, a few
journals adopt a ``by default'' hostile attitude towards preprints,
mostly due to the lack of clear policy of the publishers, or perhaps
because a preprint culture has not developed in biology and the
practice is still considered unusual. As an example, Wiley-Blackwell,
which publishes some leading journals, has no official policy on the
subject \ref{table:policies}.

@@

% Discuss embargoes?
%A second potential problem is embargoes. Many publishers subject their article
%to embargoes \cite{rea12}.

@@

% Not sure about this paragraph, as much as I love to hate MS Office, I would
% rather keep the thing as a "pro-preprint" paper only and not use it to promote
% LaTeX/R.
% -- PhDP

% JJA: It's not crucial.  My main point was that there are certain
% non-obvious cultural reasons that arXiv submissions are common.
% It's not just because arXiv started in theoretical physics.  It's
% also that people who use TeX tend to have a certain attitude, I
% would argue inherited from Don Knuth, and people who use TeX and
% LaTeX _talk to each other_.  So perhaps this was a polite way of
% saying that people in math and physics are more community-oriented,
% and people in biology tend to walk around paranoid about getting
% scooped.  Some people also might think that they can't submit to
% arXiv because they use Word.

%There are other barriers to adoption of preprint servers by biologists.  The
%most notable preprint server, arXiv, is designed to work best with \LaTeX, the
%dominant document preparation system among physicists and mathematicians.
%Jackson \cite{jackson2002preprints} argues that \LaTeX introduced an
%open source mindset to its users, who now freely share their research
%findings as well as their software.  Many biologists instead use proprietary
%software to prepare their research findings, and many journals officially prefer
%Microsoft Word documents as submissions.  The recent discipline-wide adoption of
%free software packages such as R and associated document generation systems
%\cite{xie2012}, as well as interest in open access publishing, has coincided
%with recent rise in use of preprint servers by biologists, supporting Jackson's
%claims \cite{xie12}.

@@

arXiv (\url{http://arxiv.org/}) is the most widely-used preprint server today,
and its use is almost universal in some branches of mathematics and physics.
% Ni!!!!  I really don't like starting sentences with "it"
arXiv provides a reliable citation system for all eprints and is
especially popular in high-energy physics. Physicist Paul Ginsparg
created arXiv in 1991 for theoretical high-energy physicists to
communicate preprints via email and ftp, and soon thereafter adopted
the newly created world-wide web\cite{jackson2002preprints}.  ArXiv
now receives over 7,000 submissions per month
(\url{http://arxiv.org/show_monthly_submissions}) and divides its
submissions into subcategories of physics, mathematics, computer
science, quantitative biology, finance and statistics.  The
quantitative biology category includes subcategories for Populations
and Evolution, Quantitative Methods and other categories that may be
of interest to biologists.

@@

% To review: 
Authors must either have their arXiv submission available under an
open license or grant arXiv a non-exclusive and irrevocable license to
distribute the work. In either case, arXiv does not require copyright
transfer and only requires the rights to distribute submitted articles
in perpetuity. Thus, submitting to arXiv does not in itself prevent
the authors from transferring their rights to a publisher, which is
why most publishers tolerate arXiv, even though they ask the authors
to transfer their right to them upon acceptance of the article.

@@

PeerJ (\href{https://peerj.com/}{https://peerj.com/}) is a new
publishing system that provides both a preprint server, and a peer
reviewed journal.  It is focused on the the biological and medical
sciences, which may help overcome the perception that preprints do not
have a home in biology.  PeerJ allows commenting on posted preprints,
improving the potential for pre-publication dialog. In addition,
preprints can be made private if the authors choose, and shared only
with selected colleagues. While this reduces some of the benefits of
preprints described above, it may allow some researchers who would not
otherwise post preprints to begin to explore the possibility in a
manner appropriate to their current circumstances.

@@

@@

This manuscript was developed entirely as an open project on
github. github is one of several hosting services for collaborative
development using the git revision control software.  git is a
decentralized revision control system created by Linus Torvalds used
primarily to develop software, including the Linux kernel. It allows
the users to share code, track changes, and merge versions. For
example, during the development of this manuscript, one author would
fork the project (that is: make a personal copy), modify it, and then
the changes were merged back into the main branch. These techniques
are widely used in open source software development. It goes even
farther than submitting preprints by opening the entire writing
process.

@@

Open preprint servers offer a great opportunity for open science,
especially if the community embraces the idea of discussing
preprints. Initiatives like Haldane's Sieve
(\href{http://haldanessieve.org/}{http://haldanessieve.org/}), a new
blog discussing arXiv papers in population genetics, will help make
arXiv attractive for scientists looking to promote their work. These
initiatives are important to fully exploit the potential of open
preprints servers. Posting preprints online increases the community of
available informal peer reviewers, and uses the internet for its
original community-building purposes.  Preprint servers also
facilitate communication between disciplines, bridging cultural as
well as geographic divides. The advantages are clear and the costs are
low.

@@

PDP is funded by an Alexander Graham Bell scholarship from the National Sciences
and Engineering Council of Canada.

DG is funded by a Discovery Grand from the National Sciences and Engineering
Council of Canada and by the Canada Research Chair program.

JJA is supported by NSF DEB-0614166 and NSF DEB-0919018.

@@

...

>>>>

@@

Preprints gained popularity 20 years ago with the advent of arXiv, an open
preprint server widely used in physics and mathematics \cite{gin11}. Preprints
are also integral to the culture of other scientific fields.  Paul Krugman noted
that, in economics, the \emph{traditional model of submit, get refereed,
publish, and then people will read your work broke down a long time ago. In
fact, it had more or less fallen apart by the early 80s} \cite{kru12}. In
addition to a section in arXiv, economists have also the RePEc (Research Papers
in Economics) initiative, which aims to create an archive of working papers,
manuscripts, and book chapters.  Despite the success of this approach in other
fields, most manuscripts in biology are not submitted to a public archive before
being submitted to peer-review.  In this article, we will first highlight the
advantages of open preprint servers for both scientists and publishers. We will
also debunk a few misconceptions, discuss the policies of major publishers in
biology, and briefly review the most popular open preprint servers.

\section{The case for public preprints}

The first and most often discussed advantage of arXiv and open preprints is
speed (Figure~\ref{fig:map}). The time between submission and the official
publication of a manuscript can be measured in months, sometime in years. For
all this time, the research is known only to a select few: colleagues, editors,
reviewers.  Thus, the science cannot be used, discussed, or reviewed by the
wider scientific community.  This stage of manuscript preparation is a problem
for both scientists and publishers.  Manuscripts that are unknown cannot be used
and thus take more time to be cited. It has been shown that high-energy physics,
with its high arXiv submission rate, had the highest immediacy among physics and
mathematics \cite{pra05}. Immediacy measures how quickly articles are cited.

@@

Public preprint servers offer a much fairer way to establish intellectual
priority by making the work available when done. Some manuscripts will spend
much more time in the review process than others. Surprisingly, there is a
perception in biology that public preprints make it easier to steal ideas, as if
a scientific idea was only complete once published in a peer-reviewed journal
\cite{gin11}. Ironically, mathematicians and physicists have embraced arXiv in
part to establish priority in a fair way \cite{gin11,cal12}.

@@

Submitting to a preprint server is becoming more common in biology, even though
it still involves a minority of papers. The quantitative biology section in
arXiv is experiencing faster growth in submissions than any other field
\cite{cal12}. Most scientific journals are preprint-friendly: Nature, PLOS, BMC,
PNAS, Science (mostly) \ref{table:policies}, and all the journals from Elsevier
and Springer.  The Ecological Society of America and the Genetics Society of
America recently changed their policies to allow public preprints.  Few
scientific publications will not consider a manuscript submitted to a public
preprint server.  Still, a few journals adopt a ``by default'' hostile attitude
towards preprints, mostly due to the lack of clear policy of the publishers, or
perhaps because a preprint culture has not developed in biology and the practice
is still considered unusual. As an example, Wiley-Blackwell, which publishes
some leading journals, has no official policy on the subject
\ref{table:policies}.

@@

@@

@@

arXiv (\url{http://arxiv.org/}) is the most widely-used preprint server today,
and its use is almost universal in some branches of mathematics and physics.
arXiv provides a reliable citation system for all eprints and is especially
popular in high-energy physics. Physicist Paul Ginsparg created arXiv in 1991
for theoretical high-energy physicists to communicate preprints via email and
ftp, and soon thereafter adopted the newly created world-wide
web\cite{jackson2002preprints}.  arXiv now receives over 7 000 submissions per
month (\url{http://arxiv.org/show_monthly_submissions}) and divides its
submissions into subcategories of physics, mathematics, computer science,
quantitative biology, finance and statistics.  The quantitative biology category
includes subcategories for Populations and Evolution, Quantitative Methods and
other categories that may be of interest to biologists.

@@

Authors must either have their arXiv submission available under an open license
or grant arXiv a non-exclusive and irrevocable license to distribute the work.
In either case, arXiv does not require copyright transfer and only requires the
rights to distribute submitted articles in perpetuity. Thus, submitting to arXiv
does not in itself prevent the authors from transferring their rights to a
publisher, which is why most publishers tolerate arXiv, even though they ask the
authors to transfer their right to them upon acceptance of the article.

@@

PeerJ (\href{https://peerj.com/}{https://peerj.com/}) is a new publishing system
that provides both a preprint server, and a peer reviewed journal.  It is
focused on the biological and medical sciences, which may help overcome the
perception that preprints do not have a home in biology.  PeerJ allows
commenting on posted preprints, improving the potential for pre-publication
dialog. In addition, preprints can be made private if the authors choose, and
shared only with selected colleagues. While this reduces some of the benefits of
preprints described above, it may allow some researchers who would not otherwise
post preprints to begin to explore the possibility in a manner appropriate to
their current circumstances.

@@

PeerJ uses the CC-BY-SA 3.0 license, which is similar to the CC-BY license used
by Figshare but adds the \emph{Share-alike} (sa) restriction that derivate works
needs a license identical to the license that governs the original work.

@@

This manuscript was developed entirely as an open project on github. github is
one of several hosting services for collaborative development using the git
revision control software.  git is a decentralized revision control system
created by Linus Torvalds and is used primarily to develop software, including
the Linux kernel. It allows the users to share code, track changes, and merge
versions. For example, during the development of this manuscript, one author
would fork the project (i.e.: make a personal copy), modify it, and then the
changes were merged back into the main branch. These techniques are widely used
in open source software development. It goes even farther than submitting
preprints by opening the entire writing process.

@@

Open preprint servers offer a great opportunity for open science, especially if
the community embraces the idea of discussing preprints. Initiatives like
Haldane's Sieve (\href{http://haldanessieve.org/}{http://haldanessieve.org/}), a
new blog discussing arXiv papers in population genetics, will help make arXiv
attractive for scientists looking to promote their work. These initiatives are
important to fully exploit the potential of open preprints servers. Posting
preprints online increases the community of available informal peer reviewers,
and uses the internet for its original community-building purposes.

Preprint servers also facilitate communication between disciplines, bridging
cultural as well as geographic divides. Examples include a recent series of
papers on the theory of natural selection that was posted to arXiv
simultaneously with its publication in the \emph{Journal of Evolutionary
Biology} \cite{JEB:JEB2431,JEB:JEB2498,JEB:JEB2378,JEB:JEB2373}. Other
submissions in this category include evolutionary and ecological theory by
authors trained in physics and computer science.  Since authors in these fields
regularly check arXiv, submitting preprints may be the most effective way
biologists can help others avoid ``repeated work'' \cite{de2011contribution} and
form a synthetic community of evolutionary theorists from disparate backgrounds.
The advantages are clear and the costs are low.

@@

PDP is supported by an Alexander Graham Bell scholarship from the National
Sciences and Engineering Council of Canada. JJA is supported by NSF DEB-0614166
and NSF DEB-0919018. DG is funded by a Discovery Grand from the National
Sciences and Engineering Council of Canada and by the Canada Research Chair
program.

@@

We thank Carl Boettiger for helpful comments on an earlier version of this
manuscript.
